MusicGen: A Variational Approach to Symbolic Music Generation

### 2025-05-06 paper/paper.tex
Update paper.tex

\title{Conference Paper Title\\}

\author{
\IEEEauthorblockN{Anusha Adarakati}
\IEEEauthorblockA{
\textit{KLE Technological University} \\
Hubli, India \\
01fe22bcs186@kletech.ac.in}
\and
\IEEEauthorblockN{Anirudh Dambal}
\IEEEauthorblockA{
\textit{KLE Technological University} \\
Hubli, India \\
01fe22bcs171@kletech.ac.in}
\and
\IEEEauthorblockN{Pavan Bhakta}
\IEEEauthorblockA{
\textit{KLE Technological University} \\
Hubli, India \\
01fe22bcs175@kletech.ac.in}
\and
\IEEEauthorblockN{Harish Patil}
\IEEEauthorblockA{
\textit{KLE Technological University} \\
Hubli, India \\
01fe22bcs173@kletech.ac.in}
\and
\IEEEauthorblockN{Professor Heisenburger}
\IEEEauthorblockA{
\textit{KLE Technological University} \\
Hubli, India \\
nigga@kletech.ac.in}
}

@@

>>>>

\title{MusicNetic\\}

\author{
\IEEEauthorblockN{Anusha Adarakati}
\IEEEauthorblockA{
\textit{KLE Technological University} \\
Hubli, India \\
01fe22bcs186@kletech.ac.in}
\and
\IEEEauthorblockN{Anirudh Dambal}
\IEEEauthorblockA{
\textit{KLE Technological University} \\
Hubli, India \\
01fe22bcs171@kletech.ac.in}
\and
\IEEEauthorblockN{Pavan Bhakta}
\IEEEauthorblockA{
\textit{KLE Technological University} \\
Hubli, India \\
01fe22bcs175@kletech.ac.in}
\and
\IEEEauthorblockN{Harish Patil}
\IEEEauthorblockA{
\textit{KLE Technological University} \\
Hubli, India \\
01fe22bcs173@kletech.ac.in}
\and
\IEEEauthorblockN{Professor Sharada Krishnan}
\IEEEauthorblockA{
\textit{KLE Technological University} \\
Hubli, India \\
sharada@kletech.ac.in}
\and
\IEEEauthorblockN{Professor Sharada Krishnan 2}
\IEEEauthorblockA{
\textit{KLE Technological University} \\
Hubli, India \\
sharada2@kletech.ac.in}
}

@@

\subsection{LSTM Architecture}
\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{LSTM.png}
    \caption{The Long Short-Term Memory (LSTM) network processes the piano roll sequences as time-series data.}
    \label{fig:vae_arch}
\end{figure}

 At each timestep $t$, the LSTM cell updates its hidden state $h_t$ and cell state $c_t$ through gated operations:

\begin{equation}
\begin{aligned}
f_t &= \sigma(W_f [h_{t-1}, x_t] + b_f) \\
i_t &= \sigma(W_i [h_{t-1}, x_t] + b_i) \\
o_t &= \sigma(W_o [h_{t-1}, x_t] + b_o) \\
\tilde{c}_t &= \tanh(W_c [h_{t-1}, x_t] + b_c) \\
c_t &= f_t \odot c_{t-1} + i_t \odot \tilde{c}_t \\
h_t &= o_t \odot \tanh(c_t)
\end{aligned}
\end{equation}

where $f_t$, $i_t$, and $o_t$ are the forget, input, and output gates, respectively, and $\odot$ denotes element-wise multiplication. 
\```

### LSTM Training Algorithm:

```latex
\subsection{LSTM Implementation}
The LSTM network consists of 3 recurrent layers with 512 units each, followed by a dense output layer with sigmoid activation. We implement teacher forcing with 50% probability during training and use greedy sampling for generation:

\begin{algorithm}[H]
\caption{LSTM Training}
\begin{algorithmic}[1]
\STATE Initialize hidden state $h_0$, cell state $c_0$
\FOR{each sequence $X = (x_1, x_2, \dots, x_T)$}
    \FOR{$t = 1$ to $T$}
        \STATE Predict $\hat{x}_t = \text{LSTM}(x_{t-1}, h_{t-1}, c_{t-1})$
        \STATE Update hidden state $h_t$, cell state $c_t$ using Equation (5)
        \STATE Calculate step loss $L_t = \text{BCE}(x_t, \hat{x}_t)$
    \ENDFOR
    \STATE Backpropagate through time: $\sum_{t=1}^{T} L_t$
\ENDFOR
\end{algorithmic}
\end{algorithm}
\begin{table}[h]
\caption{Model Configurations}
\label{tab:config}
\centering
\begin{tabular}{l|ll}
& VAE & LSTM \\
\hline
Latent dim & 256 & - \\
Hidden units & - & 512 $\times$ 3 \\
Loss & $\mathcal{L}_{\text{recon}} + \beta \mathcal{L}_{\text{KL}}$ & $\sum \text{BCE}$ \\
Batch size & 32 & 64 \\
Learning rate & 1e-4 & 3e-4 \\
\end{tabular}
\end{table}
